---
title: "UDF vs plain R - bfast"
author: "peterjames.zellner@eurac.edu"
date: "17/03/2022"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Libraries

```{r libraries, echo=TRUE, message=FALSE, warning=FALSE}
library("stars")
library("sf")
library("bfast")
library("mapview")
library("dplyr")
library("ggplot2")
library("units")
library("lubridate")

```

## Get NDVI Data from openEO platform

This is a NDVI time series processed on openEO cloud using [this process graph](%22~/git_projects/openeo-udf-python-to-r/r4openeo_uc2_ndvi_mskd.R%22)

```{r load_ndvi, warning = FALSE, message = FALSE}
pth_ndvi = "~/git_projects/openeo-udf-python-to-r/r4openeo_uc2_ndvi_mskd_small.nc"
ndvi = stars::read_ncdf(pth_ndvi)
st_crs(ndvi) = st_crs(32632)
```

## bfast spatial

This is the function used in py2. It is trimmed to xarray data structure, so not working in plain R.

```{r udf_fun, eval = FALSE}
source(file = "~/git_projects/openeo-udf-python-to-r/udfs/reduce_bfast.R")
```

This is the function we based the udf upon.

```{r spatial_bfm}
spatial_bfm = function(pixels, dates, start_monitor = 2018, level = c(0.001, 0.001),
                       val = "breakpoint") {
  # error handling
  stopifnot(length(pixels) == length(dates))
  stopifnot(val %in% c("breakpoint", "magnitude"))

  # create ts object for bfast
  lsts = bfastts(pixels, dates, type = c("irregular"))

  # make sure there are enough observations
  if (sum(!is.na(lsts)) < 100){
    return(NA)
  }

  # run bfast and return the selected value into the raster
  res = bfastmonitor(lsts,
                     start_monitor,
                     formula = response~harmon,
                     order = 1,
                     history = "all",
                     level = level,
                     verbose = F)[[val]]
  if(is.na(res)){
    return(0)
  }
  return(res)
}
```

## Apply bfast

Run the function on the NDVI time series in plain R.

```{r run_bfast}
level = c(0.001, 0.001)
start_monitor = 2018
dates = st_get_dimension_values(ndvi, "t")

a = Sys.time()
bfast_brks = st_apply(X = ndvi, MARGIN = c("x", "y"), function(x){
  spatial_bfm(pixels = x, 
              dates = dates, 
              start_monitor = start_monitor, 
              level = level, val = "breakpoint")
})
b = Sys.time()
duration = b-a

write_stars(obj = bfast_brks, dsn = "~/git_projects/openeo-udf-python-to-r/result_bfast_r_small.nc")
```

## Compare plain R and UDF results

### Runtime

The runtime of the UDF is for

-   50x50x400: 40 sec

-   256x256x400: 30 min

The runtime of the plain R for dims(x, y, t) `r dim(ndvi)` is `r format(duration)` 


### Spatial Coherence

Get the UDF result

```{r get_udf_res, warning = FALSE}
pth_udf_result = "~/git_projects/openeo-udf-python-to-r/result_bfast_udf_small.nc"
bfast_brks_udf = stars::read_stars(pth_udf_result)
st_crs(bfast_brks_udf) = st_crs(32632)
bfast_brks_udf[[1]][bfast_brks_udf[[1]] < 2018] = NA
bfast_brks_udf[[1]][bfast_brks_udf[[1]] > 2020] = NA
```

Get the plain R result

```{r get_r_res, warning = FALSE}
pth_r_result = "~/git_projects/openeo-udf-python-to-r/result_bfast_r_small.nc"
bfast_brks_r = stars::read_stars(pth_r_result)
bfast_brks_r[[1]][bfast_brks_r[[1]] < 2018] = NA
bfast_brks_r[[1]][bfast_brks_r[[1]] > 2020] = NA
```

Get the results from the project ECO4Alps. There we were using the Harmonized Landsat Sentinel Collection at 30m resolution.

```{r get_eco_res, warning = FALSE}
#pth_eco4alps = "/mnt/CEPH_PROJECTS/ECO4Alps/Forest_Disturbances/03_results/bfast/magn_2016_2020_start_2018_level_0.001.tif"
#bfast_brks_eco = stars::read_stars(pth_eco4alps)
#bfast_brks_eco[[1]][bfast_brks_eco[[1]] > -0.2] = NA
pth_eco4alps = "/mnt/CEPH_PROJECTS/ECO4Alps/Forest_Disturbances/03_results/bfast/brks_2016_2020_start_2018_level_0.001.tif"
bfast_brks_eco = stars::read_stars(pth_eco4alps)
bfast_brks_eco[[1]][bfast_brks_eco[[1]] < 2018] = NA
bfast_brks_eco[[1]][bfast_brks_eco[[1]] > 2020] = NA
bfast_brks_eco = st_warp(src = bfast_brks_eco, dest = bfast_brks_udf)
```

Plain R and UDF results are identical

```{r udf_vs_r}
summary(round(c(bfast_brks_r[[1]] - bfast_brks_udf[[1]]), 3)) # to the 3rd decimal
```

Plot the results

```{r mapview_results, echo = FALSE}
at_brks = seq(2018, 2020, by = 0.5)
mv_comp = mapview(bfast_brks_eco, at = at_brks) + 
  mapview(bfast_brks_r, at = at_brks) + 
  mapview(bfast_brks_udf, at = at_brks)

mv_comp
```

### Temporal Trajectory

Select a point interactively.

```{r select_point}
#pnt = mapedit::drawFeatures(map = mv_comp) %>% st_transform(st_crs(ndvi))
pnt = st_as_sf(data.frame(x = 697653.4, y = 5148280), coords = c("x", "y"), crs = st_crs(32632))
```

Extract the ndvi time series at the point.

```{r extract_ndvi}
ndvi_ts = ndvi[pnt] %>% pull() %>% c()
ndvi_ts = data.frame(ndvi = ndvi_ts, 
                     ndvi_int = zoo::na.approx(ndvi_ts),
                     dates = st_get_dimension_values(ndvi, "t"))
```

Extract the breakpoints from the bfast results.

```{r extract_brks}
brk_r = bfast_brks_r[pnt] %>% pull() %>% c() %>% date_decimal()
brk_udf = bfast_brks_udf[pnt] %>% pull() %>% c() %>% date_decimal()
brk_eco = bfast_brks_eco[pnt] %>% pull() %>% c() %>% date_decimal()
```

Plot the ndvi time series including the detected breaks.

```{r plot_ts, echo=FALSE, warning = FALSE}
brk_ts = ggplot(ndvi_ts, aes(x=dates, y=ndvi)) +
  geom_point() + 
  geom_line() +
  geom_line(aes(x = dates, y = ndvi_int), col = "gray", lty = "dashed", alpha = 0.5) + 
  geom_vline(aes(xintercept = brk_udf, col = "udf"), alpha = 0.5) + 
  geom_vline(aes(xintercept = brk_r, col = "r"), alpha = 0.5) + 
  geom_vline(aes(xintercept = brk_eco, col = "eco4alps"), alpha = 0.5)

brk_ts
```
